Typed Functional Genetic Programming

### 2013-07-10 tex/thesis.tex
[tex]

\newenvironment{todo}
{ ~\\[1em]
  \textbf{TODO}
  \begin{easylist}[itemize]}
{\end{easylist}~\\[1em]}

@@

\section{Genetic Programming}

@@

@@

\subsection{Building blocks}

@@

Satisfying this constrain ensures that every program tree build 
from \TuF will be total.\\

@@

\subsection{Generating individual}

@@

\subsection{Tree representations of \lterms}
\label{tree-reps}

@@

@@

Notice that this representation does not cover all \lterms, 
e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.  
\\\\
\Lets now consider representation for \textit{typed \lterms}.
Straightforward approach would be to add to each node a type entry which 
would be the type of the \lterm corresponding to subtree having this
node as the root node.

@@

Now we can finally put the pieces together. 
Every solution to a Inhabitation tree has this just described tree form of a typed \lterm.

\begin{todo}
& EXAMPLES of tree representations of \lterms  
& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
		Things about $\beta\eta^{-1}$ normal form, etc.   
& DESCRIBE Inhabitation Machine... 
\end{todo}

@@

@@

\begin{todo}
& Dřív se to menovalo 
"Our approach to \textit{solving} Inhabitation tree"
tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
termy a že je  budeme generovat pomocí 
řešení inhabitačních stromu.
\end{todo}

The main design choice behind our approach to solving inhabitation trees
is ability to be variable enough to enable choice between 
\textit{systematic} way of generating terms 
and \textit{classical-GP-like} way of generating terms 
by choice of simple search strategy.\\

\Lets explain what is meant by \textit{systematic} and \textit{classical-GP-like}. \\

\textit{Systematic} way of solving inhabitation tree corresponds to
generating terms in order from smallest to largest in number of symbols and \lheads.
That means minimizing number of and-nodes in and-or-tree view of point on inhabitation tree.
But since inhabitation tree is a special kind of and-or-tree with alternating layers of 
and-nodes (symbols and \lheads) and or-nodes (types) it is also minimizing number of 
size of whole inhabitation/and-or tree. \\

\textit{Classical-GP-like}

\begin{todo}
!!!
\end{todo}

@@

@@

\begin{todo}
& říct že na tu sme se moc nesoustředili ze dvou duvodu
  ktery to ospravedlňujou 
&& koza taky
&& ta s vygereovanim podstromu je v zásadě použití
   algoritmu na generovaní jedince a tak nám přišlo
   lepší pořádně se soustředit na generování jedince a
   začít do toho mutaci tahat až bude generování pořádně
   prozkoumáno.
\end{todo}

@@

@@

\section{Roadmap}
\section{Tree ants}
\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}

>>>>

\newenvironment{todo}
{ ~\\[0.5em]
  \textbf{TODO}
  \begin{easylist}[itemize]}
{ \end{easylist}
  ~}

@@

\section{Genetic Programming}
\label{GP}

@@

\begin{todo}
 & Asi to přejmenvat na S-expression.  
\end{todo}

@@

\subsection{Building blocks}
\label{building-blocks}

@@

Satisfying this constrain ensures that every program tree build 
from \TuF will be total.

We will refer to this constrain as \textit{over-one-type} constrain
in the further text.\\

@@

\subsection{Generating individual}
\label{GPgene}

@@

\section{Tree representations of \lterms}
\label{tree-reps}

\newcommand{\sexprTree}{sexpr-tree\xspace}
\newcommand{\atTree}{@-tree\xspace}

@@

We will refer to this representation as to \textit{\atTree}.\\

@@

Notice that this representation does not cover all \lterms, 
e.g. $(\lambda x.x) y$ is not expressible in it. But it does not bother us.

We will refer to this representation as to \textit{\sexprTree}.\\

\Lets now consider representation for \textit{typed \lterms}.
Straightforward approach would be to add to each node a type entry which 
would be the type of the \lterm corresponding to subtree having this
node as the root node.

@@

\begin{todo}
 & říct že to je přesně ta notace co používá koza
 & EXAMPLES of tree representations of \lterms  
 & napsat transformace mezi \atTree a \sexprTree
 & nějak to sjednotit s tim uplně na začátku v popisu GP kde se popisuje
   přímo \sexprTree a říct proč se tomu tak řiká 
   (na tom místě co je v textu dřív)

\end{todo}

\section{ Generating terms in Long normal form by solving Inhabitation tree }

Now we can finally put the pieces together. 
Every solution to an Inhabitation tree has this just described tree form of a typed \lterm.

\begin{todo}
& !!!
& TALK (more?) ABOUT "Barendregt-like" subsection \ref{barlike}\\
		Things about $\beta\eta^{-1}$ normal form, etc.   
\end{todo}

\section{ Inhabitation machine }

\begin{todo}
  & DESCRIBE Inhabitation Machine...
   && něco jako na závěrještě ve stručnosti popišme 
      IM, strukturu ze který je odvozenej IT. 
  & říct, že je to hezkej konstrukt vzhledem k tomu že je konečnej
  & ale je popsanej velmi zrychla (v tý knize) a implementační
     detaily jsou celkem vynechaný
  & na stromový verzi jsou implementační detaily hezky vidět
  & a navíc pro polymorfní verzi jsou už i IM nekonečný takže
     se strácí i ta konečnostní výhoda
  & dát tam obrázek 
\end{todo}

@@

In this chapter we will describe our system trying to 
dive into implementation details as little as possible.
Some of those details will be revealed in the next chapter
devoted to the implementation.

This chapter will be
presented in the mix between the mathematical terminology 
described in the previous chapter and the algorithmic terminology.\\

We can summarize our approach to design of 
the system for Genetic programming with types
by its main design goals:\\

\begin{enum}
 & It should eliminate the \textit{over-one-type} constrain of classical GP
   (see \ref{building-blocks}). 
 & It should be generalization of the classical GP described in \ref{GP},
   by which is meant that for \textit{building blocks} satisfying 
   constrains of classical GP there should be simple setup of control parameters
   of the system which will make the system behave in precisely the same way
   as the classical GP behaves. 
 & It should utilize the theory around typed $\lambda$-calculus.  
\end{enum}

\section{Individuals and Building blocks}

\begin{todo}
 & několik fází, několik reprezentací
 & během generování se reprezentace mění 
   && nejdří obsahuje lambdy i volný proměnný pak už ne
      &&& vygeneruje se v \sexprTree
      &&& ta se převedena na \atTree aby jsme provedli abstraction elimination
 & dvě možnosti jak uchovávat jedince po vygenerování
   && \sexprTree
      &&& currently implemented
   && \atTree
      &&& currently not implemented, ale neni problém dodělat
      &&& DOĎELAT! :)
   && ale de mezi nima volně převádět
      &&& díky tomu že nemaj lambda abstrakce, jinak by museli bejt ve 
          WHNF/HNF nebo něco takovýho
   && dusledky pro křížení
      &&& \atTree umožňuje víc křížení než \sexprTree
          &&&& každý co de v \sexprTree umí i \atTree ale obráceně to neplatí
      &&& \atTree je zase ekvivalentní tomu kozovskýmu  
 & říct že bilding blocks je prostě kontext
\end{todo}

@@

%\begin{todo}
%& Dřív se to menovalo 
%"Our approach to \textit{solving} Inhabitation tree"
%tak tu napsat pěknej uvod o tom, že naši jedinci jsou lambda
%termy a že je  budeme generovat pomocí 
%řešení inhabitačních stromu.
%\end{todo}

Input for term generating algorithm is pair $(\sigma,\Gamma)$ where:\\

\begin{enum}
 & $\sigma$ is desired type of each generated individual.
 & $\Gamma$ is \textit{building blocks} context. 
\end{enum}~

Each generated \lterms $M$ should  satisfy \GMS. \\

We generate terms in long normal form.
This is achieved by \textit{solving} inhabitation tree for $(\sigma,\Gamma)$.  
The main design choice behind our approach to solving inhabitation trees
is ability to be variable enough to enable choice between 
\textit{systematic}  
and \textit{ramped half-and-half} (described in \ref{GPgene}) 
methods of generating terms 
by choice of simple search strategy.\\

\Lets explain what is meant by \textit{systematic} 
method of generating terms.

Systematic way of solving inhabitation tree corresponds to
generating terms in their long normal forms
in order from smallest to largest in number of symbols and \lheads.
That means minimizing number of and-nodes in and-or-tree view of point on inhabitation tree.
But since inhabitation tree is a special kind of and-or-tree with alternating layers of 
and-nodes (symbols and \lheads) and or-nodes (types) it is also minimizing number of 
size of whole inhabitation/and-or tree.

\begin{todo}
 & Ten poslední odstavec je napsanej divně, přepsat.
   && říct v něm že na to budem teda používat A*. 
\end{todo}

@@

\subsection{Other tried approach to generating individuals}

\begin{todo}
 & popsat tu ten puvodní přístup, kde se bralo přímo 
   zavedení a eliminace pro každej typovej konstrukt 
   a podle toho se pak udělali příslušný term generating grammar pravidla
   kerý odpovídaj manipulacímse strommem
 & ale má nevýhody
   && hlavně to neni v normální formě, takže generujeme spoustu 
      ruzně dopočítanejch beta-rovnáse-ekvivalentních termu
   && u pravidla zavedení implikace neni uplně zřejmý
      co tam použít za typ
   && takže sice to na to ze začátku běhalo ale nakonec to bylo vyměněno
      za IM metodu since ta je jednoduší, líp se přemejšlí o ní,
      konstrukty jako and a or nemusíme zavádět na urovni pravidel
      ale stačí tam dat vhodný polymorfní konstruktory/destruktory 
 & nějak tu metodu pomenovat a přemenovat podle toho nadpis
\end{todo}

@@

\begin{todo}
& říct že na tu sme se moc nesoustředili ze dvou duvodu
  ktery to ospravedlňujou 
	&& koza taky
	&& ta s vygereovanim podstromu je v zásadě použití
	   algoritmu na generovaní jedince a tak nám přišlo
	   lepší pořádně se soustředit na generování jedince a
	   začít do toho mutaci tahat až bude generování pořádně
	   prozkoumáno.
& To ale neznamená že jí nemáme radi, naopak se těšíme
  že její vhodný zavedení celou metodu ještě zlepší.
\end{todo}

@@

\section{ Complexity discussion }

\begin{todo}
& !!!
& zmínit že mi moc neřešíme formální/asymptotický složitosti
  && de nám hlavně o realný časy
& vyjít z komplexnosti A* asi
  && složitost výrazně ovlivněná kvalitou heuristický funkce
\end{todo}

@@

\section{Roadmap}
\section{Tree ants}
\section{[Propojení těch dvou barandrechtskejch odvozovacích pravidel do jedinýho]}
\section{[Chytřejší heuristika pro A* která si to předpočítá na $\overline{\Gamma}$ ]}
\section{[šlechtění těch strategií na generování termů]}
\section{[šlechtění fitness funkcí a nápad s "turnajem olympioniků"]}
